---
title: "Daily Assignment 21"
author: "Avery Eastman"
format: html
execute: 
  echo: true
---

```{r}
library(dataRetrieval)
library(dplyr)
library(tsibble)


# Example: Cache la Poudre River at Mouth (USGS site 06752260)
poudre_flow <- readNWISdv(siteNumber = "06752260",    # Download data from USGS for site 06752260
                          parameterCd = "00060",      # Parameter code 00060 = discharge in cfs)
                          startDate = "2013-01-01",   # Set the start date
                          endDate = "2023-12-31") |>  # Set the end date
  renameNWISColumns() |>                              # Rename columns to standard names (e.g., "Flow", "Date")
  mutate(Date = yearmonth(Date)) |>                   # Convert daily Date values into a year-month format (e.g., "2023 Jan")
  group_by(Date) |>                                   # Group the data by the new monthly Date
  summarise(Flow = mean(Flow))                       # Calculate the average daily flow for each month
```

# Covert to tsibble

### Use as_tsibble() to convert the data.frame into a tsibble object. This will allow you to use the feast functions for time series analysis
```{r}
poudre_tbl <- as_tsibble(poudre_flow)
head(poudre_tbl)
```

# Plotting the time series

### Use ggplot to plot the time series data. Animate this plot with plotly
```{r}
library(ggplot2)
library(plotly)
poudre_plot <- ggplot(poudre_tbl, aes(x = Date, y = Flow)) +
  geom_line(color = "steelblue") +
  labs(title = "Monthly Streamflow of Cache la Poudre River",
       x = "Date", y = "Average Monthly Flow") +
  theme_minimal()

ggplotly(poudre_plot)
```

# Subseries

### Use gg_subseries to visualize the seasonal patterns in the data. This will help you identify any trends or seasonal cycles in the streamflow data.
```{r}
library(feasts)

gg_subseries(poudre_tbl) +
  labs(title = "Monthly Streamflow",
       x = "Year", y = "Flow") +
  theme_minimal()
```

### Describe what you see in the plot. How are “seasons” defined in this plot? What do you think the “subseries” represent?
The plot shows a seasonal pattern in streamflow for the Cache la Poudre River. You can see that streamflow peaks in May-June, which can be classified as late spring and early summer. This is most likely due to snow melt as the weather gets warmer in these months. You can also see that streamflow is lowest in November-Febuary, which can be classified as late fall and winter. 

# Decompose

### Use the model(STL(...)) pattern to decompose the time series data into its components: trend, seasonality, and residuals. Chose a window that you feel is most appropriate to this data…
```{r}
poudre_decomp <- poudre_tbl |>
  model(STL(Flow ~ season(window = 13))) |>
  components()

autoplot(poudre_decomp)
```

### Describe what you see in the plot. How do the components change over time? What do you think the trend and seasonal components represent?
The trend component increases from 2013-2015ish, then from 2016-2020 decreases, and after 2020 remains fairly stable. This trend component shows the fluctuation in streamflow over time, which could be reflective of changes in watershed hydrology. The seasonality component shows repeated peaks and valleys throughout the years. The peaks for this component are most likely during spring since there is likely be more snowmelt during that time (creating more streamflow). The remainder component varies a lot over the time period in the plot. 
